---
title: "MFA 2017"
author: "Nick Browen"
date: "August 22, 2018"
output: html_document
---

Data
```{r}
#Soil <- read.csv("C:/Users/nickbrowen/Desktop/STAT Summer Research/Soil2018.csv")[,-1]
#Soil <- data.frame(Soil_data[,1], row.names = Soil_data$Order)

Bud <- read.csv("C:/Users/nickbrowen/Desktop/STAT Summer Research/MarSoil.csv")[,-c(1,11)]
Blm <- read.csv("C:/Users/nickbrowen/Desktop/STAT Summer Research/MaySoil.csv")[,-c(1,11)]
Ver <- read.csv("C:/Users/nickbrowen/Desktop/STAT Summer Research/JulSoil.csv")[,-c(1,11)]
Har <- read.csv("C:/Users/nickbrowen/Desktop/STAT Summer Research/SepSoil.csv")[,-c(1,11)]
```

```{r}
names(Bud)[-1] <-  paste("Bud", names(Bud), sep = "_")[-1]
names(Blm)[-1] <-  paste("Blm", names(Blm), sep = "_")[-1]
names(Ver)[-1] <-  paste("Ver", names(Ver), sep = "_")[-1]
names(Har)[-1] <-  paste("Har", names(Har), sep = "_")[-1]
```

Merging
```{r}
temp.Bud.Blm <- merge(Bud, Blm, all = T, by = "Order")
temp.Bud.Blm.Ver <- merge(temp.Bud.Blm, Ver, all = T, by = "Order")
Soil <- merge(temp.Bud.Blm.Ver, Har, all = T, by = "Order")
Soil <- data.frame(Soil, row.names = Soil$Order)[-1]
Soil[is.na(Soil)] <- 0
```

Packages
```{r}
library(FactoMineR)
library(factoextra)
library(ggsci)
```


MFA
```{r}
res.mfa <- MFA(Soil, group = c(8,8,8,8), type = rep("c", 4), name.group = c("Budbreak", "Bloom", "Veraison", "Harvest"), graph = F)
```

Results
```{r}
eigenval <- res.mfa$eig
fviz_screeplot(res.mfa, addlabels = T)
```

Standardized Loadings (Correlation) for Groups
```{r}
load.season <- sweep(res.mfa$group$coord,2,sqrt(res.mfa$eig[1:ncol(res.mfa$group$coord),1]),FUN="/")
load.season
```


```{r}
group <- get_mfa_var(res.mfa, "group")
fviz_mfa_var(res.mfa, "group", title = "Season Coordinates")
```

Contributions of Seasons
```{r}
fviz_contrib(res.mfa, "group", axes = 1)
fviz_contrib(res.mfa, "group", axes = 2)
```

Variables
```{r}
quanti.var <- get_mfa_var(res.mfa, "quanti.var")
fviz_mfa_var(res.mfa, "quanti.var", palette = "jco", 
             col.var.sup = "violet", repel = TRUE,  geom = c("point", "text"), legend = "bottom")
```

```{r}
fviz_mfa_var(res.mfa, "quanti.var", col.var = "contrib", repel = TRUE, 
             geom = c("point", "text"))
```


```{r}
fviz_mfa_var(res.mfa, choice = "quanti.var", axes = c(1,2))
```


Contributions of Variables
```{r}
fviz_contrib(res.mfa, choice = "quanti.var", axes = 1, top = 23,
             palette = "jco")
fviz_contrib(res.mfa, choice = "quanti.var", axes = 2, top = 23,
             palette = "jco")

```

Partial Axes
```{r}
fviz_mfa_axes(res.mfa)
```

```{r}
load.partial.axes <- sweep(res.mfa$partial.axes$coord,2,sqrt(res.mfa$eig[1:ncol(res.mfa$partial.axes$coord),1]),FUN="/")
load.partial.axes
```


```{r}
#Partial Axes (each seasons dimensions) correlation with MFA dim1
 sort(load.partial.axes[,1][load.partial.axes[,1] >0.35 | load.partial.axes[,1] < -.35])
#all partial axes are season's dim1
```


```{r}
#Partial Axes (each seasons dimensions) correlation with MFA dim2
sort(load.partial.axes[,2][load.partial.axes[,2] >0.35 | load.partial.axes[,2] < -.35])
```


```{r}
#Correlation of Season's dim1 with MFA's dim2
res.mfa$partial.axes$cor[,2][c("Dim1.Budbreak", "Dim1.Bloom", "Dim1.Veraison", "Dim1.Harvest")]
```


Loadings for MFA Dim1
```{r}
#sweep(res.mfa$global.pca$var$coord,2,sqrt(res.mfa$global.pca$eig[1:ncol(res.mfa$global.pca$var$coord),1]),FUN="/")
```


Variables by MFA Dim1 and Dim2
```{r}
fviz_mfa_var(res.mfa)
```

Variables most correlated with MFA dim1 and dim2 **could be the weirdly high corrs
```{r}
#dim1 (all of them)
sort(res.mfa$quanti.var$cor[,1][res.mfa$quanti.var$cor[,1] >0.35 | res.mfa$quanti.var$cor[,1] < -.35])

#dim2
 sort(res.mfa$quanti.var$cor[,2][res.mfa$quanti.var$cor[,2] >0.35 | res.mfa$quanti.var$cor[,2] < -.35])
```




Will be useful for how specific microbes change over time
```{r}
fviz_mfa_ind(res.mfa, partial = c("sphingobacteriales", "actinomycetales", "bacillales", 
+ "solirubrobacterales"))
```

```{r}
Soil[121,] #sphingobacteriales
Soil[7,] #actinomycetales
```


```
For sphingobacteriales, this order has high number of microbes in soil microbial communities with non-CC treatments in Bloom and Budbreak, for Veraison and Harvest CC treatments are still lower number of microbes than the other treatments but not by that much, basically the non-CC treatments are pretty low during Harvest and Ver (hence low Dim2 coord), but the non-CC trts are still a little higher than the CC trts (hence slightly higher Dim1 coord).
```
```{r}
fviz_contrib(res.mfa, choice = "ind", axes = c(1,2), top = 23,
             palette = "jco")
```



MFA DIM1
```{r}
dim1contrib <- data.frame(Cont = res.mfa$quanti.var$contrib[,1], Trt = row.names(res.mfa$quanti.var$contrib), Group = rep(c("H", "NH", "SF", "OF", "NF", "HWD", "LWD", "NCC"), 4))

dim1contrib$Trt <- factor(dim1contrib$Trt, levels = dim1contrib$Trt[order(-dim1contrib$Cont)])

ggplot(dim1contrib, aes(x = Trt, y= Cont, fill = Group)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + geom_hline(yintercept =  100/32, linetype = "dotted") +  ggtitle("Variable Contributions to MFA Dim1 Colored by Treatment Level")
```


```{r}
trt_dim1contrib <- data.frame(Cont = res.mfa$quanti.var$contrib[,1], Trt = row.names(res.mfa$quanti.var$contrib), Group = rep(c("Herb", "Herb", "Fert", "Fert", "Fert", "CC", "CC", "CC"), 4))

trt_dim1contrib$Trt <- factor(trt_dim1contrib$Trt, levels = trt_dim1contrib$Trt[order(-trt_dim1contrib$Cont)])

ggplot(trt_dim1contrib, aes(x = Trt, y= Cont, fill = Group)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + geom_hline(yintercept =  100/32, linetype = "dotted") + ggtitle("Variable Contributions to MFA Dim1 Colored by Treatmeant Type")
```


```{r}
season_dim1contrib <- data.frame(Cont = res.mfa$quanti.var$contrib[,1], Trt = row.names(res.mfa$quanti.var$contrib), Group = rep(c("Budbreak", "Bloom", "Veraison", "Harvest"), each = 8))

season_dim1contrib$Trt <- factor(season_dim1contrib$Trt, levels = season_dim1contrib$Trt[order(-season_dim1contrib$Cont)])

ggplot(season_dim1contrib, aes(x = Trt, y= Cont, fill = Group)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + geom_hline(yintercept =  100/32, linetype = "dotted") + ggtitle("Variable Contributions to MFA Dim1 Colored by Season")
```






MFA DIM2
```{r}
dim2contrib <- data.frame(Cont = res.mfa$quanti.var$contrib[,2], Trt = row.names(res.mfa$quanti.var$contrib), Group = rep(c("H", "NH", "SF", "OF", "NF", "HWD", "LWD", "NCC"), 4))

dim2contrib$Trt <- factor(dim2contrib$Trt, levels = dim2contrib$Trt[order(-dim2contrib$Cont)])

ggplot(dim2contrib, aes(x = Trt, y= Cont, fill = Group)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + geom_hline(yintercept =  100/32, linetype = "dotted") +  ggtitle("Variable Contributions to MFA Dim2 Colored by Treatment Level")
```

```{r}
trt_dim2contrib <- data.frame(Cont = res.mfa$quanti.var$contrib[,2], Trt = row.names(res.mfa$quanti.var$contrib), Group = rep(c("Herb", "Herb", "Fert", "Fert", "Fert", "CC", "CC", "CC"), 4))

trt_dim2contrib$Trt <- factor(trt_dim2contrib$Trt, levels = trt_dim2contrib$Trt[order(-trt_dim2contrib$Cont)])

ggplot(trt_dim2contrib, aes(x = Trt, y= Cont, fill = Group)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + geom_hline(yintercept =  100/32, linetype = "dotted") + ggtitle("Variable Contributions to MFA Dim2 Colored by Treatmeant Type")
```

```{r}
corr_dim2contrib <- data.frame(Cont = res.mfa$quanti.var$contrib[,2], Trt = row.names(res.mfa$quanti.var$contrib), Group = rep(c("H", "NH", "SF", "OF", "NF", "HWD", "LWD", "NCC"), 4), Corr = res.mfa$quanti.var$cor[,2], Pos_Corr = res.mfa$quanti.var$cor[,2] > 0)

corr_dim2contrib$Trt <- factor(corr_dim2contrib$Trt, levels = corr_dim2contrib$Trt[order(-corr_dim2contrib$Cont)])

ggplot(corr_dim2contrib, aes(x = Trt, y= Cont, fill = Pos_Corr)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + geom_hline(yintercept =  100/32, linetype = "dotted") +  ggtitle("Variable Contributions to MFA Dim2 Colored by Correlation")
```

